### 2023-03-02 reports/paper/01-introduction.rmd
manuscript prepared

Non-Hodgkin Lymphoma (NHL) is the seventh most common cancer type in the United States and has garnered intense research attention because its incidence in the United States doubled between 1973 and 1994, before plateauing at around 19 per 100,000 persons [@seer_1994;@Ekstrom-Smedby_2006]. The strongest known risk factor of NHL is immunosuppression, both congenital and acquired [@Filipovich_1992;@Chiu_2015]. Modulation of immune function contributes to the pathogenesis of NHL directly by perturbing lymphocyte function or indirectly by reducing immunosurveillance against neoplasms. Immunodeficiency may also interact with infection with Epstein-Barr Virus, resulting in uncontrolled B-cell proliferation [@Oertel_2002]. Acquired immunodeficiency through infection with human immunodeficiency virus (HIV) can lead to genetic lesions and dysregulation while increasing vulnerability to infection by other oncogenic agents [@Grulich_1999;@Alexander_2007]. However, since immunodeficiency and infection with HIV are rare, they cannot fully explain the historic rise or present burden of NHL [@Shiels_2013]. Since the rise in NHL incidence coincided with a period of rapid and extensive chemicalization in industry, agriculture and warfare, there has been stronger interest in the environmental and occupational epidemiology on NHL [@Nelson_2005;@Romero_2021]. A meta-analysis of 44 articles published between 1980 and 2014 found statistically significant associations between NHL and exposure to pesticides of several classes including carbamate, organophosporus, triazine, and organochlorine [@Schinasi_2014]. While this meta-analysis presents striking evidence of the potential explanatory role of occupational exposures, it also highlights the limited breadth of occupational exposures studied.

Studies of NHL and other important occupational exposures including oils, solvents, and lubricants remain inconclusive [@Ekstrom-Smedby_2006]. Furthermore, since NHL is relatively rare, nearly all studies of NHL incidence follow a single time-point case-control design [@Schinasi_2014;@Alexander_2007]. Large multicenter studies including EPILYMPH and InterLymph attained large enough numbers of cases to investigate particular NHL subtypes, but despite their impressive size and comprehensiveness, were limited by the lack of quantitative time-varying exposure information and the lack of adjustment for time-varying sources of bias [@Besson_2006;@Morton_2014]. In the case of occupational studies, the healthy worker survivor effect (HWSE) is a key concern. The HWSE is the dynamic selection process by which healthier individuals remain at work, where they accumulate more occupational exposure, and less healthy individuals leave work [@Arrighi_1994]. Standard analyses of occupational cohorts affected by the HWSE result in a downward bias on the exposure-outcome associations of interest, as less healthy workers typically accrue less exposure than healthier workers.

Metalworking fluids (MWF), an important class of occupational exposures, are complex mixtures of oils and chemicals used to lubricate and cool metal during manufacturing processes. During use, MWFs aerosolize and present a potential health hazard to exposed workers. The type of MWF applied depends in part on the particular metalworking operation. Straight MWF, the oldest class of MWF, is comprised mostly of mineral oil. Concerns over the carcinogenicity of MWF began as early as the 1970s, when the International Agency for Research on Cancer (IARC) classified mineral oils as carcinogenic, citing studies of occupational exposures among workers in oil, textile, and metal industries [@IARC_1973]. The combination of health concerns, the high cost of oil in the 1970s, and the introduction of performance-enhancing water-soluble chemical additives led to the gradual replacement of straight MWF with water-based soluble and synthetic MWFs for many, but not all operations [@Childers_2005]. The elimination of particular etiologic agents often leads to their replacement with other potentially hazardous chemical compounds, however [@Kassotis_2020;@Maertens_2021]. Although water-based MWF contain little to no mineral oil, they contain a richer cocktail of additives including chromates, cycloalkanes, phenols, organochlorines, nitrites, sulfonates, and triazines: classes of chemicals with known or plausible carcinogenicity [@IARC_1987;@IARC_2014;@IARC_2017]. Hence, limits on entire classes of exposures grouped by functional use may result in stronger, more lasting protections for worker health than limits on particular etiologic agents because they will remain in place even as the composition and formulation of complex occupational exposures change [@Maertens_2021]. More than 20 years after the IARC classification of mineral oils as carcinogenic, the National Institute for Occupational Safety and Health (NIOSH) concluded that there exists substantial evidence linking all MWF exposure to several different cancers and now recommends an exposure limit of 0.5 mg/m^3^ for total particulate matter derived from any type of MWF [@niosh_1998;@Mirer_2003].

Here, we assess the effect of hypothetical MWF exposure limits on NHL cumulative incidence in the United Auto Workers-General Motors (UAW-GM) Occupational Cohort Study. In particular, we assess the effect of limits on soluble MWF that (1) ignore exposure to other types of MWF and (2) account for other types of MWF exposure by intervening more strongly on soluble MWF when exposure to all types of MWF is high. The latter approach is more consistent with the NIOSH REL, which was recommended for exposure to any type of MWF. The large size of the study population and rich time-varying, quantitative MWF exposure data provide an opportunity to study this relatively rare cancer and evaluate realistic interventions on MWF exposure in a longitudinal cohort setting. A recent study of cancer incidence in this iconic occupational cohort found a monotonic association between NHL and exposure to soluble MWF [@Colbeth_2022]. This association was based on a traditional Cox proportional hazards regression, a method vulnerable to bias due to confounding by time-varying factors affected by prior exposure. We therefore explore this association using a novel causal estimator capable of adjusting for time-varying confounding and censoring affected by past exposure. This estimator also yields marginal effect estimates of hypothetical interventions, which are more compatible with a population health framework than conditional measures.

>>>>

Non-Hodgkin Lymphoma (NHL) is the seventh most common cancer type in the United States and has garnered substantial research attention because its incidence in the United States doubled between 1973 and 1994, before plateauing at around 19 per 100,000 persons per year [@seer_1994;@Ekstrom-Smedby_2006]. The strongest known risk factor of NHL is immunosuppression, both congenital and acquired [@Filipovich_1992;@Chiu_2015]. Modulation of immune function contributes to the pathogenesis of NHL directly by perturbing lymphocyte function or indirectly by reducing immunosurveillance against neoplasms. Immunodeficiency may also interact with infection with Epstein-Barr Virus, resulting in uncontrolled B-cell proliferation [@Oertel_2002]. Acquired immunodeficiency through infection with human immunodeficiency virus (HIV) can lead to genetic lesions and dysregulation while increasing vulnerability to infection by other oncogenic agents [@Grulich_1999;@Alexander_2007]. However, since immunodeficiency and infection with HIV are rare, they cannot fully explain the historic rise or present burden of NHL [@Shiels_2013]. Since the rise in NHL incidence coincided with a period of rapid and extensive chemicalization in industry, agriculture and warfare, there has been stronger interest in the environmental and occupational epidemiology on NHL [@Nelson_2005;@Romero_2021].

Pesticide exposure in agricultural occupational settings was a common target of NHL research in recent decades. A meta-analysis of 44 articles published between 1980 and 2014 found statistically significant associations between NHL and exposure to pesticides of several classes including carbamate, organophosporus, triazine, and organochlorine [@Schinasi_2014]. Occupations associated with NHL are not limited to the agricultural sector, however. Metal processors, health workers, salespeople, machinists, and electricians have all been identified as occupational groups with higher risk of NLH [@Fritschi_2005a; @Mester_2006; @Ekstrom-Smedby_2006]. Studies contrasting occupational groups to the general population are vulnerable to exposure misclassification of the true causal agent. When different causal agents contribute to overall NHL risk by their effect on particular subtypes of NHL, their etiologic role may be obscured in analyses not stratified by subtype. Large multicenter studies including EPILYMPH and INTERLYMPH seek to disentangle effect heterogeneity across NHL subtypes, but are limited by the lack of quantitative time-varying exposure information [@Besson_2006;@t-Mannetje_2008;@Morton_2014].

Measures of time-varying exposure are essential to statistical adjustment for forms of time-varying confounding affected by past exposure such as the healthy worker survivor effect (HWSE). The HWSE is the dynamic selection process by which healthier individuals remain at work, where they accumulate more occupational exposure, and less healthy individuals leave work [@Arrighi_1994]. Standard analyses of occupational cohorts affected by the HWSE result in a downward bias on the exposure-outcome associations of interest, as less healthy workers typically accrue less exposure than healthier workers.

Metalworking fluids (MWF), an important class of occupational exposures, are complex mixtures of oils and chemicals used to lubricate and cool metal during manufacturing processes. During use, MWFs aerosolize and present a potential health hazard to exposed workers. The type of MWF applied depends in part on the particular metalworking operation. Straight MWF, the oldest class of MWF, is comprised mostly of mineral oils, which were classified as carcinogenic by the International Agency for Research on Cancer (IARC) in 1973, citing studies of occupational exposures among workers in oil, textile, and metal industries [@IARC_1973]. The combination of health concerns, the high cost of oil in the 1970s, and the introduction of performance-enhancing water-soluble chemical additives led to the gradual replacement of straight MWF with water-based soluble and synthetic MWFs for many, but not all operations [@Childers_2005].

Although water-based MWF contain little to no mineral oil, they contain a richer cocktail of additives including chromates, cycloalkanes, phenols, organochlorines, nitrites, sulfonates, and triazines: classes of chemicals with known or plausible carcinogenicity [@IARC_1987;@IARC_2014;@IARC_2017]. The elimination of particular etiologic agents often leads to their replacement with other potentially hazardous chemical compounds [@Kassotis_2020;@Maertens_2021]. Hence, limits on entire classes of exposures grouped by functional use may result in stronger, more lasting protections for worker health than limits on particular etiologic agents because they will remain in place even as the composition and formulation of complex occupational exposures change [@Maertens_2021]. More than 20 years after the IARC classification of mineral oils as carcinogenic, the National Institute for Occupational Safety and Health (NIOSH) concluded that there exists substantial evidence linking all MWF exposure to several different cancers. However, the NIOSH recommended exposure limit of 0.5 mg/m^3^ for total particulate matter derived from any type of MWF was intended to reduce respiratory disorders and not cancer [@niosh_1998;@Mirer_2003].

Here, we assess the effect of hypothetical MWF exposure limits on NHL cumulative incidence (which we will refer to as risk) from 1985 to 2005 in the United Auto Workers-General Motors (UAW-GM) Occupational Cohort Study. The large size of the study population and rich time-varying, quantitative MWF exposure data provide an opportunity to study this relatively rare cancer and evaluate realistic interventions on MWF exposure in a longitudinal cohort setting. A recent study of cancer incidence in this iconic occupational cohort found a monotonic association between NHL and exposure to soluble MWF [@Colbeth_2022]. This association was based on standard Cox proportional hazards regression, a method vulnerable to bias due to confounding by time-varying factors affected by past exposure. We therefore explore this association using an estimator capable of adjusting for time-varying confounding and censoring affected by past exposure: the hazard-extended iterative conditional expectation (ICE) parametric g-formula. Like other causal estimators in statistics, the hazard-extended ICE g-formula yields marginal effect estimates of hypothetical interventions, which are more compatible with a population health framework than conditional measures.

The g-formula is a well-known result in causal inference for the identification of causal effects in the presence of time-varying confounding affected by past exposure [@Robins_1986]. Standard representations of the g-formula include (1) a non-iterated expectation over the joint density of covariates, (2) the ICE over time, and (3) an inverse probability weighted expectation. The parametric g-formula is a plug in estimator for the g-formula under its first, non-iterative, representation. It involves the parametric modeling of the full joint distribution of the outcome, exposure, and covariates under study for each time point [@Taubman_2009;@Young_2014]. Counterfactual quantities under hypothetical interventions of interest are computed from Monte Carlo samples from distributions implied by the fitted parametric models. In longitudinal settings, this approach often requires specifying and fitting large number of models in order to satisfy the exchangeability assumptions necessary for causal identification.
<!-- Since applications of the parametric g-formula are often context-laden, there are no unified guidelines for model specification. -->

When researchers are not wholly comfortable with the full set of parametric assumptions required by the parametric g-formula, the ICE g-formula may be preferred. Parametric estimators using the ICE representation of the g-formula require modeling only the conditional outcome distributions, so they require fewer parametric assumptions than non-iterative parametric g-formula estimators. Counterfactual outcome estimates over the follow-up period are computed from interval-specific conditional estimates by applying the tower rule of expectation. Under the assumptions of conditional exchangeability at all time points, positivity, counterfactual consistency, and correct model specification, the hazard-extended parametric g-formula yields unbiased estimates of counterfactual risk with greater statistical efficiency than both propensity score-based estimators and the classical ICE g-formula.

### 2023-03-02 reports/paper/02-methods.rmd
manuscript prepared

We estimated NHL risk from 1985 to 2005 under hypothetical limits on average annual exposure to soluble MWF by applying the novel hazard-extended iterative conditional expectation (ICE) parametric g-formula estimator [@Wen_2020]. We leveraged time-varying quantitative MWF exposure data in tandem with employment records to adjust for the HWSE. First, we estimated the expected number of NHL cases per 1000 workers that we would observe if there were no competing risks. We contrasted  this counterfactual risk to that when in addition, the hypothetical exposure limits of 0.5, 0.25, and 0.05 mg/m^3^ were enforced for soluble MWF over workers' entire working lifetimes. These hypothetical limits are static interventions because they do not consider other covariates or exposures. However, since workers were often exposed to multiple types of MWF, both daily and cumulatively, we also evaluated the effect of limiting total MWF exposure to those hypothetical limits by intervening only on exposure to soluble MWF. These interventions, which take into consideration exposure to MWF of other types when deciding how much soluble MWF exposure to assign, are dynamic interventions on exposure to soluble MWF. All of the hypothetical interventions we evaluate are stochastic interventions rather than deterministic interventions because post-intervention exposures take on a range of values.

@@

We identified incident cancers in the UAW-GM Cohort that occurred between 1985 and 2004 by linkage to the Michigan Cancer Registry (MCR). Workers at Plants 1 and 2, located in the greater Detroit metropolitan area, were also linked to the Detroit Regional Registry of the Surveillance, Epidemiology, and End Results (SEER) Program. Cancer types were distinguished using site and histology codes conforming to the International classification of Diseases for Oncology, 3^rd^ edition (ICD-O-3). Non-Hodgkin lymphoma was defined by cancers with any of the following ICD-O-3 Histology codes: 9590-9597, 9670-9671, 9673, 9675, 9678-9680, 9684, 9687-9691, 9695, 9698-9702, 9705, 9708-9709, 9712, 9714-9719, 9724-9729, 9735, 9737-9738, 9811-9818, 9823, 9827, 9837. Details regarding cancer incidence follow-up are described elsewhere [@Colbeth_2022]. Vital status was ascertained from company records and by linkage to Social Security Administration, National Death Index, and state mortality files.

Covariates including year of hire, sex, race, and plant location were obtained from company records. Race was missing for about 16% of the cohort, most commonly among workers hired before 1960 in Plant 2. In analyses, missing race was considered a distinct category.

@@

Company industrial hygienists collected several hundred personal and area samples for total particulate matter (mg/m^3^) composed of MWF over many decades. Research industrial hygienists collected additional air sampling data when the cohort study was launched in the mid 1980s. These data were combined with the historical data to derive quantitative 8-hour time-weighted average exposure estimates to soluble, straight, and synthetic MWFs for each combination of job, department, and plant over time. Workers' time-weighted average annual exposure to each MWF type was determined by combining this job-exposure matrix with employment records, which recorded time-varying job type, department, and plant through 1994. For employment records that were at least half complete, gaps in the record were interpolated by carrying forward the last known job type. The exposure assessment is described in detail elsewhere [@Hallock_1994;@Woskie_1994;@Woskie_2003]. Previous analyses of NHL applied exposure lags of 1 to 20 years; we lagged cumulative MWF exposures by 10 years to account for disease latency and therefore ended follow-up on January 1, 2005 [@Smith_2007;@Karipidis_2007;@Zhang_2019].

@@

We applied the hazard-extended ICE parametric g-formula with pooling over treatment history [@Wen_2020] to estimate 20-year counterfactual risk of non-Hodgkin lymphoma. We considered six stochastic interventions on average annual exposure to soluble MWFs after the elimination of censoring/competing risks. We considered hypothetical exposure limits starting with the NIOSH Recommended Exposure Limit (REL) for total particulate mass (PM) from MWFs (0.5 mg/m^3^). Then, we estimated the risk at half the REL (0.25 mg/m^3^), and a tenth of the REL (0.05 mg/m^3^) [@niosh_1998]. The first three interventions were static interventions which enforced these hypothetical limits on average annual exposure to soluble MWF. The remaining three interventions were dynamic interventions that reduced average annual exposure to soluble MWF with the goal of enforcing the hypothetical limits for total exposure to MWF, regardless of type. 
Under these interventions, exposures in some person-years did not comply with the hypothetical limit on total exposure to MWF. That is, if the sum of exposures to straight and synthetic MWF exceeded the hypothetical limit, average annual exposure to soluble MWF was simply reduced to 0; we did not evaluate interventions on exposures to the other two classes of MWF.
Interventions were applied at hire, before the start of follow-up, through the end of follow-up. Person-years were not intervened upon if the observed average annual exposure was below the hypothetical limit. Figure 1 illustrates the static and dynamic interventions for three person-year examples. Panel A shows the observed levels of exposure to soluble MWF and to MWF of other types. Panel B shows the post-intervention levels of exposure under a static intervention capping exposure to soluble MWF at the hypothetical limit (dotted line). Panel C shows the post-intervention levels under a dynamic intervention reducing exposure to soluble MWF with the goal of capping total MWF exposure at the hypothetical limit. Note that in person-year 3, the observed level of exposure to total MWF (and thus, to soluble MWF) is below the hypothetical limit, so the post-intervention exposure is left at the observed level.

![Figure 1. Observed and post-intervention levels of exposure to soluble MWF and MWF of other types in three example person-years. The white boxes represent exposure to soluble MWF, and the shaded boxes represent exposure to MWF of other types. The horizontal dotted line represents the hypothetical exposure limit.](`r here::here("resources/images", "interventions.png")`){width=6.5in}

The hazard-extended ICE parametric g-formula estimator is an ICE expression of the parametric g-formula estimator further extended to incorporate the predicted hazard of past outcomes [@Wen_2020]. The classical ICE g-formula estimator is algebraically equivalent to the noniterative conditional expectation (NICE) g-formula estimator, which typically involves computing covariate-standardized estimates sequentially, starting at the index time, by specifying parametric models for the joint distribution of the confounders, exposures, and outcomes at each time point [@Taubman_2009]. In contrast, ICE g-formula estimators avoid the need for modeling the entire joint distribution of observed variables and only require outcome models for each time point. Counterfactual outcome estimates over the follow-up period are computed from interval-specific conditional estimates by applying the tower rule of expectation. Under the assumptions of conditional exchangeability at all time points, positivity, counterfactual consistency, and correct model specification, the hazard-extended parametric g-formula yields unbiased estimates of counterfactual risk with greater statistical efficiency than both propensity score-based estimators and the classical ICE g-formula.

Our implementation of the hazard-extended ICE parametric g-formula involved a series of model-based standardization steps using logistic regression. Post-intervention estimates of the discrete hazard of NHL given all exposures and covariates were combined iteratively from the end of follow-up to the start. In each iteration, predicted discrete hazards were standardized over post-intervention exposure and covariate histories before combining with discrete hazards from the previous iteration. This iterative process resulted in sequentially standardized estimates of NHL risk over the entire follow-up period. Averaging over the baseline distribution of covariates yielded the counterfactual risk estimate of NHL when the intervention of interest was enforced for the entire study population.

We split the 20-year follow-up period into eight time periods; the first two periods spanned four years each, and the remaining six periods spanned two years each. The first two periods are longer in length to account for the smaller number of cases in those years. Post-intervention exposure and exposure history were summarized as cumulative exposure. We modeled discrete hazards by fitting a pooled logistic regression for NHL over at-risk person-periods given cumulative exposure to straight, soluble, and synthetic MWFs, employment status, cumulative time off, year of hire, sex (male/female), race (Black/white/unknown), and plant (Plant 1/Plant 2/Plant 3). Cumulative exposure to MWFs, employment status, and cumulative time off were lagged 10 years. All continuous variables were represented as categorical variables with cut points determined by the tertiles of nonzero values among NHL cases. During the iterative combination of discrete hazards, we performed model-based standardization over baseline covariates and the complete set of time-varying covariate histories.

We estimated risk under the observed distribution of soluble MWF exposure (natural course) and under the six interventions. We contrasted the risk under intervention to that under the natural course by computing relative risks. Confidence intervals were computed using the nonparametric bootstrap with 1000 samples and centering on the estimate computed from observed data.

<!-- GITHUB? -->

>>>>

We estimated NHL risk from 1985 to 2005 under hypothetical limits on average annual exposure to soluble MWF by applying the hazard-extended iterative conditional expectation (ICE) parametric g-formula estimator [@Wen_2020]. We leveraged time-varying quantitative MWF exposure data in tandem with employment records to adjust for the HWSE. First, we estimated the expected number of NHL cases that we would observe if there were no censoring by competing risks. Then, we contrasted this counterfactual risk to that when in addition, the hypothetical exposure limits of (1) the NIOSH Recommended Exposure Limit (REL) for total particulate mass (PM) from MWFs (0.5 mg/m^3^), (2) half the REL (0.25 mg/m^3^), and (3) a tenth of the REL (0.05 mg/m^3^) were enforced for soluble MWF over workers' entire working lifetimes [@niosh_1998]. Under these hypothetical interventions, exposure was set at the hypothetical limit if observed exposure was greater than that limit.  Otherwise, exposure not intervened upon. This kind of intervention is known as a threshold longitudinal modified treatment policy [@Taubman_2009;@Richardson_2013;@Haneuse_2013;@Young_2014]. Longitudinal modified treatment policies are intervention strategies that depend on the natural value of exposure at a particular time ie the value of exposure that would have been observed at that time if the intervention were discontinued immediately prior [@Richardson_2013;@Diaz_2021].

@@

We identified incident cancers in the UAW-GM Cohort that occurred between 1985 and 2005 by linkage to the Michigan Cancer Registry (MCR). Workers at Plants 1 and 2, located in the greater Detroit metropolitan area, were also linked to the Detroit Regional Registry of the Surveillance, Epidemiology, and End Results (SEER) Program. Cancer types were distinguished using site and histology codes conforming to the International classification of Diseases for Oncology, 3^rd^ edition (ICD-O-3). Non-Hodgkin lymphoma was defined by cancers with any of the following ICD-O-3 Histology codes: 9590-9597, 9670-9671, 9673, 9675, 9678-9680, 9684, 9687-9691, 9695, 9698-9702, 9705, 9708-9709, 9712, 9714-9719, 9724-9729, 9735, 9737-9738, 9811-9818, 9823, 9827, 9837. Subtypes of NHL were also identified for sensitivity analyses. Details regarding cancer incidence follow-up are described elsewhere [@Colbeth_2022]. Vital status was ascertained from company records and by linkage to Social Security Administration, National Death Index, and state mortality files.

Covariates including year of hire, sex, race, and plant location were obtained from company records. Race was missing for about 16% of the cohort, most commonly among workers hired before 1960 in Plant 2. In analyses, missing race was considered a distinct category. All covariates were coded as categorical variables. Cut-points for continuous covariates were determined according to the quantiles among cases.

@@

Quantitative measurement of time-varying MWF exposure is a distinguishing strength of the UAW-GM cohort relative to other studies of occupational exposure. Exposure assessment was based on direct sampling as well as company records. Company industrial hygienists collected several hundred personal and area samples for total particulate matter (mg/m^3^) composed of MWF over many decades. Research industrial hygienists collected additional air sampling data when the cohort study was launched in the mid 1980s. These data were combined with the historical data and company records to derive quantitative 8-hour time-weighted average exposure estimates to soluble, straight, and synthetic MWFs for each combination of job, department, and plant over time. Workers' time-weighted average annual exposure to each MWF type was determined by combining this job-exposure matrix with employment records, which recorded time-varying job type, department, and plant through 1994. For employment records that were at least half complete, gaps in the record were interpolated by carrying forward the last known job type. The exposure assessment is described in detail elsewhere [@Hallock_1994;@Woskie_1994;@Woskie_2003]. Previous analyses of NHL applied exposure lags of 1 to 20 years; we lagged cumulative MWF exposures by 10 years to account for disease latency and therefore ended follow-up on January 1, 2005 [@Smith_2007;@Karipidis_2007;@Zhang_2019]. In analyses, MWF exposure history was summarized as the cumulative sum of annual exposure intensities and coded as categorical variables with cut-points determined according the quantiles of cumulative exposure among cases.

@@

The hazard-extended ICE parametric g-formula may be thought of as a series of model-based standardization steps, which we implemented using logistic regression. We split the 20-year follow-up period into eight time periods; the first two periods spanned four years each, and the remaining six periods spanned two years each. The first two periods are longer in length to account for the smaller number of cases in those years. Post-intervention estimates of the discrete hazard of NHL given all exposures and covariates were combined iteratively from the end of follow-up to the start. In each iteration, predicted discrete hazards were standardized over post-intervention exposure and covariate histories before combining with discrete hazards from the previous iteration. This iterative process results in sequentially standardized estimates of NHL risk over the entire follow-up period. Averaging over the baseline distribution of covariates yields the counterfactual risk estimate of NHL when the intervention of interest was enforced for the entire study population.

Post-intervention exposure and exposure history were summarized as cumulative exposure. We modeled discrete hazards by fitting a pooled logistic regression for NHL over at-risk person-periods given cumulative exposure to straight, soluble, and synthetic MWFs, employment status, cumulative time off, year of hire, sex (male/female), race (Black/white/unknown), and plant (Plant 1/Plant 2/Plant 3). Cumulative exposure to MWFs, employment status, and cumulative time off were lagged 10 years. All continuous variables were represented as categorical variables with cut points determined by the tertiles of nonzero values among NHL cases. An overview of the general steps of the estimation procedure are presented below.

0.  For the hypothetical exposure limit of interest, compute the cumulative exposure
    each worker would have accrued by the end of each follow-up period since hire.
1.  Fit a pooled logistic regression to the observed data for NHL on covariates and
    cumulative exposure over all at-risk person-periods, excluding those ending with a
    censoring event.
2.  Replace the cumulative exposure vector in the observed data with the
    post-intervention cumulative exposure vector. Using the model fitted in the 
    previous step, compute estimates of the post-intervention discrete hazard for
    each at-risk person-period including those that end with a censoring event.
3.  Set $k = K - 1$ where $K = 8$ is the number of follow-up periods:

	a) Among those who were event-free and uncensored through the $k$^th^ period,
		 fit a logistic regression on the predicted post-intervention discrete hazard
		 for the period spanning the $(k + 1)$ through $K$^th^ periods given observed
		 covariate and exposure history up through (and including) the $k$^th^ period.
	b) Replace the cumulative exposure vector in the observed data with the
		 post-intervention cumulative exposure vector. Using the model fitted in the
		 previous step, obtain predicted values for all those who were at-risk in the
		 $k$^th^ period.
	c) Compute the predicted post-intervention discrete hazard for the period spanning
		 the $k$ through $K$^th^ periods by multiplying the predicted values from the
		 previous step by 1 minus the discrete hazard estimate used to fit the model
		 in step 3.a and adding that same discrete hazard estimate to the product.
	d) If $k > 1$, set $k = k - 1$ and return to step 3.a.

4.  Compute the counterfactual risk by averaging the predicted post-intervention
		discrete hazards for the entire follow-up period for all units.

We estimated risk under the observed distribution of soluble MWF exposure (natural course) and under the six interventions. We contrasted the risk under intervention to that under the natural course by computing relative risks. Confidence intervals were computed using the nonparametric bootstrap with 1000 Monte Carlo samples from the population at risk at start of follow-up and centering on the estimate computed from observed data. All the necessary script used to reproduce the analyses are available on [GitHub](https://github.com/kvntchn/gm-nhl-ice).
